---
title: "Lab15"
author: "Kimberly"
date: "2/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r The Data}
# setting up the data
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)

# checking
head(counts)
head(metadata)

# how many genes are we dealing with?
nrow(counts)
```
- Important: The sample IDs in the "metadata" sheet (SRR####) (column 1) exactly match the column names of the "counts" data (column 2-8), except for column 1, which contains the Ensembl gene ID.

## Calculate avg of counts per gene for control samples and compare it to those of drug-treated samples.(See beginning of lecture notes.)
- See if there is a difference overall for a given gene.

**CONTROL**
```{r Searching for the correct dataset}
metadata$dex == "control"

# we want to access the ID of the controls
#control <- metadata[metadata$dex == "control", ]$id
# control

control <- metadata[metadata$dex == "control",]
# AKA metadata$id[ metadata$dex == "control" ]

# code below gives you table of all control data
# metadata[metadata$dex == "control", ]

counts[, control$id]
```
```{r Finding the avg of CONTROL samples}
control.mean <- rowSums(counts[ , control$id] ) / nrow(control)
control.mean
names(control.mean) <- counts$ensgene
```

**TREATED**
```{r Finding the avg of TREATED samples}
metadata$dex == "treated"

treated <- metadata[metadata$dex == "treated",]

counts[, treated$id]

treated.mean <- rowSums(counts[, treated$id]) / nrow(treated)
treated.mean
```

Combining our meancount data for bookkeeping purposes.
```{r}
meancounts <- data.frame(control.mean, treated.mean)
meancounts
```

## Creating a plot
Inspection: treated.mean vs. control.mean
- if there is no effect, data should land on a straight line (averages would also be the same point).
```{r}
plot(meancounts)
```
We had almost 40,000 genes, but these dots do not look like 40,000 points. The dense area is all the gene data bunched up.
- Take the *log of your data* to spread informaton out (and reduce skewing).
```{r}
# log = "xy" makes it so that both x and y axis are logged
plot(meancounts, log = "xy")
```
- point outside above the line = more expression in *treated*
- point outside below the line = more expression in *control*

```{r}
meancounts$log2fc <- log2( meancounts[, "treated.mean"]/meancounts[, "control.mean"] )

head(meancounts)
# focus on the last column titled "log2fc"
```
- if we have NaN, we are trying to take log of 0
- Remove the NaN values (these result from zero avg. values in .mean)
```{r}
# "which" function tells you which position in the vector fits the condition of what is True
# example
x <- data.frame(c(1, 1, 10, 0), c(0, 1, 10 , 0))
x
# T if 0, F if not 0
x == 0
# finds position of which are '0'
which(x == 0)
# finds position of which are '0' and puts it in an array
which(x == 0, arr.ind = T)


ind <- unique( which (x == 0, arr.ind = T)[, "row"])
x[-ind]
```

```{r}
zero.vals <- which(meancounts[, 1:2] == 0, arr.ind = T)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)

nrow(mycounts)
```
Common threshold convention:
if fold change (fc) is > 2, significant up change #upregulation, if < -2, significant down change #downregulation 

```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```

How many of each do I have?
```{r # of upreg; # of downreg}
sum(up.ind)
sum(down.ind)
```
- To a statistician, FC can sometimes be considered meaningless. FC can be large w/o sig. difference from t-test or ANOVA.
- To a biologist, FC is almost always considered important.
1. v small but stat sig FC might not be relevant to a cell's fxn.
2. interest to kow which genes are most dramatically regulated, as these are often thought to reflect changes in bio meaningful transcripts and/or pathways

## DESeq2 analysis
- calculates FC AND p-value

Set up input for DESeq by specifying countdata, metadata, and the design (i.e., whchih comparisons we are interested in)
```{r Set up the input}
library(DESeq2)
citation("DESeq2")
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```
```{r DESeq()}
# The DESeq() function takes a DESeqDataSet and returns a DESeqDataSet, but with lots of other information filled in
dds <- DESeq(dds)

res <- results(dds)
res
```

## Volcano plot
- We are plotting the p-values, specically
- each dot is a gene
- upregulated on the right; downregulated on the left
```{r Plotting the information}
plot(res$log2FoldChange, res$padj)
```
```{r Logging the plot}
plot(res$log2FoldChange, log(res$padj))
```
To flip the axis to make it look like a volcano, put the (-) log. (Just how the rest of the world does it.)
```{r Make it look like volcano}
plot(res$log2FoldChange, -log(res$padj))
```

Highlighting upregulated and downregulated genes, genes that are sig diff.
```{r}
mycols <- "gray"
plot(res$log2FoldChange, -log(res$padj), col = mycols)

# make a vector that is specific
mycols <- rep("gray", nrow(res))
# overwrite the gray ones that are above the change of 2-fold
mycols[ res$log2FoldChange > 2] <- "red"
mycols[ res$log2FoldChange < (-2)] <- "blue"
plot(res$log2FoldChange, -log(res$padj), col = mycols)
abline(v=c(2, -2), col = "gray", lty=2)
abline(h = log(0.05), col = "gray", lty = 2)
```

