---
title: "Class18_Lab1: Investigating cancer genomics datasets"
author: "Kimberly"
date: "3/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

NCI-GDC (cancer genomics commons) is a major portal for cancer genomics data.

Here we use the **GenomicDataCommons** bioconductor package, as it provides functions for querying, accessing, and mining the NCI-GDC in R.
- talk to this database and DL data etc.
BiocManager::install( c("GenomicDataCommons", "TCGAbiolinks", "maftools") )

```{r}
library(GenomicDataCommons)
library(TCGAbiolinks)
library(maftools)
```

Check if we can talk to GDC
```{r}
status()
# make sure the $status section says "OK"
```

## 3. Querying the GDC from R
```{r}
# %>% is a pipe operator
# can be useful in R and sometimes is more clear to read: "Here is the info, this is what we do to it."
# ex) sum( c(1:10)) == c(1:10) %>% sum()

cases_by_project <- cases() %>%
  facet("project.project_id") %>%
  aggregations()
head(cases_by_project)
```

```{r}
x <- cases_by_project$project.project_id
x
```

### Make a bar plot of the overview data
```{r}
barplot(x$doc_count)
```
Heavily skewed! Should normalize it with log transformations.

```{r}
colvec <- rep("lightblue", nrow(x))

barplot(x$doc_count, log = "y",
        names.arg = x$key, las = 2, col=colvec)
```

## 4. Variant analysis w/ R
- After some exploration of the website I came up with the following query: “cases.project.project_id in ["TCGA-PAAD"] and files.data_type in ["Masked Somatic Mutation"] and files.data_format in ["MAF"].

Find files available for PAAD project that we can then examine variants from.
```{r}
maf.file <- GDCquery_Maf(tumor="PAAD", pipelines = "mutect")
head(maf.file)
```

### MAF analysis
Read in the  MAF file we found:
```{r}
vars = read.maf(maf = maf.file, verbose = FALSE)
```

Now make a summary plot of this data
`plotmafSummary()` is a function that exists within the maftools package.
```{r}
plotmafSummary(vars)
```

### Drawing oncoplots or waterfall plots
```{r}
oncoplot(maf = vars, top = 10)
```
- What are the top mutated genes? --> 77% (all missense mutations)