---
title: "Class09_Lab"
author: "Kimberly"
date: "2/4/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

https://bioboot.github.io/bimm143_W20/class-material/lab-8-bimm143.html
```{r open the csv file}
# takes the first column and gets rid of the variable name "X"
x <- read.csv("UK_foods.csv", row.names = 1)
x
```
*Q1. How many rows and columns are in your new data frame named `x`? What R functions could you use to answer this question?*
```{r Q1}
# Complete the following code to find out how many rows and columns are in x
dim(x)
```

Let's explore the data with plots
```{r bar graph plotting pt. i}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```
```{r bar graph plotting pt. ii}
# Try setting beside=FALSE in your barplot() code.
barplot(as.matrix(x), beside= F, col=rainbow(nrow(x)))

# This plot is still not very helpful in comparison
```

*Q5: Generating all pairwise plots may help somewhat. Can you make sense of the following code and resulting figure?*
- *What does it mean if a given point lies on the diagonal for a given plot?*
```{r Q5: Pairwise plots}
pairs(x, col=rainbow(10), pch=16)
# A5: first row: English vs. Wales, England vs. Scotland, England vs. N. Ireland
# ON line on diagonal = SAME proportions of the food between each country
# helpful but still tedious
```

## Principal Component Analysis (PCA) to the rescue with `prcomp()` function
```{r creating the PCA data}
# t takes the transpose (rotates the data so foods are in cols instead)
pca <- prcomp(t(x))
pca
summary(pca)
```
To read the summary data:
- "Approximately 67% of the variance in the data is accounted for by the first principal component ("Proportion of Variance), and approximately 97% is accounted for in total by the first two principal components (Cumulative Proportion)."

Check the attributes/components of the output w/ `attributes()`
```{r Attributes of output}
attributes(pca)
```


*Q7. Complete the code below to generate a plot of PC1 vs PC2. The second line adds text labels over the data points.*
```{r Q7: Plotting PCA data}
# Plot PC1 vs PC2
plot(pca$x[, 1], pca$x[, 2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x), 
     col=c("grey", "red", "blue", "orange"))
```

