---
title: "Class13"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Download and clean HIV-Pr receptor structure
- DL 1HSG PDB structure from online database, remove bound drug from Merk and write out a protein-only and ligand-only new set of PDB files.
- We will be using `get.pdb()`, `read.pdb()`, `atom.select()`, and `write.pdb()` functions from **bio3d package**.
```{r Getting the files}
# load package (recall: you already installed it a while back inside the console)
library(bio3d)
# get. will download the data for you
file.name <- get.pdb("1hsg")

hiv <- read.pdb(file.name)
hiv
```
```{r Protein-only and Ligand-only}
# select protein only
prot <- atom.select(hiv, "protein", value = T)
lig <- atom.select(hiv, "ligand", value = T)
# can also use trim.pdb()
## prot <- trim.pdb(hiv, "protein")
## lig <- trim.pdb(hiv, "ligand")

# make the files
write.pdb(prot, file = "1hsg_protein.pdb")
write.pdb(lig, file = "1hsg_ligand.pdb")
# can view the files with bio3d.view()
# to check protein file, can print to see that there is only MK1

```
So far,
PDB
>protein
>>.pdbqt
>>>added hydrogens; atom types; charges
>ligand

# 1.5 Prepare a docking configuration file
(check the '1.5 docking config file' text file)
- This deﬁnes your protein (receptor), ligand (ligand) number of docking modes to generate (num_modes). 
- All the docked modes will be collated in a ﬁle deﬁned by out (all.pdbqt).

# Docking ligands into HIV-1 protease with Autodock Vina
- download the application
- in your terminal, use command: 
> "\Program Files (x86)\The Scripps Research Institute\Vina\vina.exe" -config config.txt --log log.txt

- you should have 2 new files all.pdbqt and log.txt

### Inspecting your docking results
```{r}
library(bio3d)
res <- read.pdb("all.pdbqt", multi = T)
write.pdb(res, "results.pdb")
```
Lower affinity will be closer to the actual.
- In VMD, you can see results.pdb molecule at mode 1 is most similar to MK1.


