---
title: "Class07"
author: "Kimberly"
date: "1/28/2020"
output: html_document
---

# echoes all the code that is ran
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# My functions
- you can modify this in the console or md and execute it, it will update your vairables in the environment tab (righthand side)
```{r}
# # Our rescale function from the end of lecture 9
rescale <- function(x, na.rm=TRUE, plot=FALSE, ...) {

  # '!' means 'not'
  if(na.rm) {
    rng <-range(x, na.rm=TRUE)
  } else {
    rng <-range(x)
  }

  answer <- (x - rng[1]) / (rng[2] - rng[1])
  if(plot) { 
    plot(answer, ...) 
  }

  return(answer)
}
```

# Creating your own error warnings
## warning() and stop()
both output text at certain times
- warning() continues on after the warning
-- use for tolerable error
- stop() breaks the function
-- use for intolerable error

```{r}
rescale2 <- function(x, na.rm=TRUE, plot=FALSE, ...) {
  # Our rescale function from lecture 10

  if( !is.numeric(x) ) {
    stop("Input x should be numeric", call.=FALSE)
  }
  
  rng <-range(x, na.rm=TRUE)

  answer <- (x - rng[1]) / (rng[2] - rng[1])
  if(plot) { 
    plot(answer, ...) 
  }

  return(answer)
}

```

# is.na() helps find missing values
- also works on character type vectors
```{r}
x <- c(1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
z <- c(NA, 2)

print(is.na(x))
# outputs [1] FALSE FALSE  TRUE FALSE  TRUE

# check if x & y have NA in the same position
print(is.na(x) & is.na(y))
# outputs [1] FALSE FALSE  TRUE FALSE FALSE

```

## Write a function to check if x & y have NA in same position
```{r}
is_both_na <- function(x, y) {
  if(length(x) != length(y)) {
    stop("Your inputs aren't the same length")
  }
  is.na(x) & is.na(y)
  
}
```

### modified both_na3 function
```{r}
rescale_new <- function(x, y) {
  ## Print some info on where NA's are as well as the number of them 
  if(length(x) != length(y)) {
    stop("Input x and y should be vectors of the same length", call.=FALSE)
  }
  na.in.both <- ( is.na(x) & is.na(y) )
  na.number <- sum(na.in.both) # T = 1, F = 0; gives u count of NAs in both x and y
  na.which <- which(na.in.both) # gives position of T (NAs)
  
  message("Found ", na.number, " NA's at position(s):", 
          paste(na.which, collapse=", ") ) 
  
  return( list(number=na.number, which=na.which) )
}
```

# DATA FRAMES
- tables basically
formatting:
- subsetting df[row, column]
-- df[, 1] <-- gives you entire first column (outputted in a single line, so it looks like a row though)
-- df[, "header_name"] <-- works the same way as above
-- df$IDs <-- works the same way as above
-- gene.colname = "header name" <-- works the same way as above
-- df[ ,gene.colname] <-- works the same way as above

```{r}
# Which genes are present in BOTH datasets/frames?
x <- df1$IDs
y <- df2$IDs

intersect(x, y)
# outputs "gene2" "gene3"

# Maybe you want the entire dataset compared? (Not just gene names)

## What are positions of x in y?
### takes all of x: which elements of x are in y (returns boolean)
x %in% y
# outputs [1] FALSE  TRUE  TRUE

# subset: x (df1 IDs) of x in y takes those that are T
x[x %in% y]
# outputs "gene2" "gene3"

```

making a matrix 
```{r}
cbind(x[x %in% y], y[y %in% x])
```
making a function using the matrix
```{r}
# highlight piece of code, >code >extract function
# modifies the highlighted code
gene_intersect <- function(x, y) {
  cbind(x[x %in% y], y[y %in% x])
}

x <- df1$IDs
y <- df2$IDs

gene_intersect(x, y)
```

# function that does everything we've done so far
```{r}

gene_intersect2 <- function(df1, df2) { 
   cbind( df1[ df1$IDs %in% df2$IDs, ], # takes T/F portion (stuff inside []) and subsets the entire df1 (IDs and exp) using those T/F results
          df2[ df2$IDs %in% df1$IDs, "exp"] )
}


gene_intersect3 <- function(df1, df2, gene.colname="IDs") { 
   cbind( df1[ df1[,gene.colname] %in% df2[,gene.colname], ], 
          exp2=df2[ df2[,gene.colname] %in% df1[,gene.colname], "exp"] )
}

# can specify diff columns with the last parameter
gene_intersect3(df1, df2, gene.colname = "IDs")

## colnames("new_name", "old_name") changes names of headers
# colnames(df1) <- c("new_names", "exp")
```

# readability
```{r}
gene_intersect4 <- function(df1, df2, gene.colname="IDs") { 

  df1.name <- df1[,gene.colname]
  df2.name <- df2[,gene.colname]

  df1.indexes <- df1.name %in% df2.name
  df2.indexes <- df2.name %in% df1.name

   cbind( df1[ df1.indexes, ], 
          exp2=df2[ df2.indexes, "exp"] )
}

gene_intersect4(df1, df2)
# gene_intersect4(df1, df2, gene.colname = "IDs") works too
```
