---
title: "Class 6 R functions"
author: "Kimberly"
date: "1/23/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# '#' headings in R markdown work the same way as in python

Let's see more about **file importing/reading**. The main read function in base R is `read.table()`.

## Importing test1.txt
```{r}
# read.table("test1.txt") will put all data in col 1
# make it right
t1 <- read.table("test1.txt", sep = ",", header = T)
t1
# could also just use read.csv("test1.txt")
```

## Importing test2.txt
```{r}
# read.table("test2.txt") puts everything in col 1 again
# fix it
read.table("test2.txt", header = T, sep = "$")
```

## Importing test3.txt
```{r}
read.table("test3.txt", header = T)
```

Writing functions:
### FORMAT
`name_of_function` <- function(arg1, arg2,...) {
  statements/conditions
  return(something)
}

### How does it work with **VECTORS** as inputs?
```{r}
add <- function(x, y = 1) {
  x + y
}

print(add(c(1,2,4)))
# foiling
print(add(c(1,2,4), c(1,2,4)))
print(add(c(1,2,4), 4))
```

## range() function
```{r}
x <- c(4, 4, 10, 3, 11)
range(x)
```

```{r}
rescale <- function(x) {
  rng <- range(x)
  (x - rng[1]) / (rng[2] - rng[1])
}
print(rescale(1:10))

# code below outputs NA NA NA NA NA
# rescale( c(1, 2, NA, 3, 10))
# gives range NA NA when calling rng <- range(x)
# FIX IT by going to function body to take into account NA
# NEW FUNCTION adds in argument 'na.rm = T'
```
```{r}
rescale_new <- function(x) {
  rng <- range(x, na.rm = T)
  (x - rng[1]) / (rng[2] - rng[1])
}
rescale_new( c(1, 2, NA, 3, 10))
```

EXTRA STUFF from lecture slides
```{r}
rescale_3 <- function(x, na.rm=TRUE, plot=FALSE) {    
  rng <-range(x, na.rm=na.rm)    
  print("Hello")   
  
  answer <-(x - rng[1]) / (rng[2] - rng[1])   
  
  print("is it me you are looking for?")   
  
  if(plot) {       
    plot(answer, typ="b", lwd=4)    
  }   
  print("I can see it in ...")
  return(answer)
}

rescale_3(x)
```

